Ainur:~A Framework for Repeatable End-to-End Wireless Edge Computing Testbed Research

### 2022-12-19 body/glossary.tex
glossary fixes

\newabbreviation{WCA}{WCA}{Wearable Cognitive Assistance}
\newabbreviation{AR}{AR}{Augmented Reality}
\newabbreviation{VR}{VR}{Virtual Reality}
\newabbreviation{XR}{XR}{eXtended Reality}
\newabbreviation{GPU}{GPU}{Graphical Processing Unit}
\newabbreviation{FPS}{FPS}{Frames Per Second}
\newabbreviation{DNN}{DNN}{Deep Neural Network}
\newabbreviation{CPS}{CPS}{Cyber-Physical System}
\newabbreviation{NCS}{NCS}{Networked Control System}
%\acroindefinite{NCS}{an}{a}

@@

\newabbreviation{NTP}{NTP}{Network Time Protocol}

@@

\newabbreviation{SDR}{SDR}{Software-Defined Radio}

@@

\newabbreviation{HTTP}{HTTP}{Hyper-Text Transfer Protocol}

@@

%\acroindefinite{NCS}{an}{a}

>>>>

\newabbreviation{WCA}{WCA}{Wearable Cognitive Assistance}
\newabbreviation{AR}{AR}{Augmented Reality}
\newabbreviation{VR}{VR}{Virtual Reality}
\newabbreviation{XR}{XR}{eXtended Reality}
\newabbreviation{GPU}{GPU}{Graphical Processing Unit}
\newabbreviation{FPS}{FPS}{Frames Per Second}
\newabbreviation{DNN}{DNN}{Deep Neural Network}
\newabbreviation{CPS}{CPS}{Cyber-Physical System}
\newabbreviation[prefixfirst={a\ }, prefix={an\ }]{NCS}{NCS}{Networked Control System}
%\acroindefinite{NCS}{an}{a}

@@

\newabbreviation[prefixfirst={a\ }, prefix={an\ }]{NTP}{NTP}{Network Time Protocol}

@@

\newabbreviation[prefixfirst={a\ }, prefix={an\ }]{SDR}{SDR}{Software-Defined Radio}

@@

\newabbreviation[prefixfirst={a\ }, prefix={an\ }]{HTTP}{HTTP}{Hyper-Text Transfer Protocol}

@@

### 2022-12-19 body/part1/introduction.tex
glossary fixes

This thesis presents three core contributions to the existing body of research in edge computing, as well as a number of secondary ones.
Firstly, we introduce a methodological approach to studying system responsiveness versus resource consumption trade-offs in edge-bound feedback-loop systems such as \gls{WCA} and \gls{NCS}.
This approach is based on the emulation of the client-side workload, while maintaining the \emph{real} server-side process as well as network stack.
We validate this methodology by example, presenting implementations of its application in the context of \gls{WCA} and \acp{NCS}, and introduce a software framework for the orchestration of the edge computing testbeds necessary for its implementation.

@@

\cref{paper:olguinmunoz2022cleave}, on the other hand, presents a discussion on the implementation of this methodology for \acp{NCS}.
As discussed above, these applications are similarly difficult as \gls{WCA} to benchmark (in particular at scale on multi-tenant systems) due to their client-side complexity, but in particular because of their often extreme sensitivity to latency.
\todo[inline]{More argumentation or refer to previous section.}
We apply thus in this work our methodology to \gls{NCS} through the implementation of a tool for the emulation and subsequent deployment of these systems on edge computing infrastructure.
This tool allows us to both emulate the physical components of a relatively simple control system plant and deploy real algorithms for its control.
The software acts as a middleware which abstracts away the network from the development of these workloads, allowing for quick prototyping and deployment.
This work also showcases the scalability and flexibility of our approach, allowing us to deploy scenarios with a large number of loops without the need for domain-specific hardware.

>>>>

This thesis presents three core contributions to the existing body of research in edge computing, as well as a number of secondary ones.
Firstly, we introduce a methodological approach to studying system responsiveness versus resource consumption trade-offs in edge-bound feedback-loop systems such as \gls{WCA} and \gls{NCS}.
This approach is based on the emulation of the client-side workload, while maintaining the \emph{real} server-side process as well as network stack.
We validate this methodology by example, presenting implementations of its application in the context of \gls{WCA} and \glspl{NCS}, and introduce a software framework for the orchestration of the edge computing testbeds necessary for its implementation.

@@

\cref{paper:olguinmunoz2022cleave}, on the other hand, presents a discussion on the implementation of this methodology for \glspl{NCS}.
As discussed above, these applications are similarly difficult as \gls{WCA} to benchmark (in particular at scale on multi-tenant systems) due to their client-side complexity, but in particular because of their often extreme sensitivity to latency.
\todo[inline]{More argumentation or refer to previous section.}
We apply thus in this work our methodology to \gls{NCS} through the implementation of a tool for the emulation and subsequent deployment of these systems on edge computing infrastructure.
This tool allows us to both emulate the physical components of a relatively simple control system plant and deploy real algorithms for its control.
The software acts as a middleware which abstracts away the network from the development of these workloads, allowing for quick prototyping and deployment.
This work also showcases the scalability and flexibility of our approach, allowing us to deploy scenarios with a large number of loops without the need for domain-specific hardware.

### 2022-12-19 body/part1/overview/methodology.tex
glossary fixes

Two main arguments drive this approach.
One, the difficulty of scaling the number of clients in these systems in a research context.
The systems we target exhibit a centralized nature in which a potentially large number of clients offload computation to a single central compute node.
The client-side component of the applications of interest for this research is often complex to scale, either due to the involvement of humans (as in the case of \gls{WCA}) or due its cyber-physical nature (in \acp{NCS}).
Emulating this component reduces this complexity by moving it into the software domain, allowing for easier scaling through the use of cheap, general-purpose hardware such as single-board computers (e.g. Raspberry Pis, Beaglebones, etc.).

@@

In the following, we will discuss the initial implementation of this methodology to two different classes of applications.
We first discuss the methodology in the context of a human-in-the-loop application, \acl{WCA}, in section \cref{ssec:methodology:wca}.
Then, in \cref{ssec:methodology:ncs}, we present an exploration into it's utility for the study of \acf{NCS}.

\subsection{Applying the methodology to \acs{WCA}}\label{ssec:methodology:wca}

@@

\subsection{Applying the methodology to \acsp{NCS}}\label{ssec:methodology:ncs}

@@

\begin{figure*}
    \centering
    \includegraphics[width=.8\textwidth]{publications/2022CLEAVE/images/CLEAVE_NCS_structure}
    \caption{
        Structure of an emulated \acl*{NCS} in \acs*{CLEAVE}.
    }\label{fig:cleave:ncs:struct}
\end{figure*}

We overcome this issue by proposing a completely virtual plant allowing for unparalleled flexibility in changing the plant model and characteristic features of the experiments.
In this paper, we present the first fully-software-based framework for scalable and repeatable benchmarking of edge-native \gls{NCS}.
As edge computing begins being adopted by industry, more and more variations have begun to appear in literature.
``Near'', ``far'', ``core'', and ``telco'' edge describe variations of the original concept and are becoming ubiquitous in new research.
While the core idea of edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strengths and weaknesses of such different edge concepts is of paramount importance.

Our framework, \gls{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems.
It is fully virtualized, inspired by our previous work on benchmarking human-in-the-loop applications on the edge~\cite{Olguin2019EdgeDroid}.
The tool consists of a benchmarking framework and software development kit for the development of emulated physical systems and softwarized controllers.
These virtual \acp{NCS} can then be deployed on real networks for reparametrizable, repeatable, and reproducible benchmarking of the complete system.

@@

We validate the utility of this tool through an example use case approximating a proposed edge deployment of inverted pendula control loops co-located with video analytics services.
We argue such a use case represents a realistic scenario and appropriate benchmark for the tool, since
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the inverted pendulum plant is ubiquitous in \gls{NCS} research
    \item similar setups exist in real-world industrial use
    \item video analytics has long been proposed as a ``killer app'' for edge computing.
\end{enumerate*}
Our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
We believe \gls{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \acp{NCS}.

@@

\begin{figure}[t]
    \centering
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_topple_frac}
        \caption{Toppled plants.}\label{fig:video:toppled}
    \end{subfigure}%
    \hfill%
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_angle_rms}
        \caption{Angle \gls{RMS}.}\label{fig:video:rms}
    \end{subfigure}\\
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_drop_frac}
        \caption{Packet losses.}\label{fig:video:drop}
    \end{subfigure}%
    \hfill%
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_rtt}
        \caption{\acsp{RTT}.}\label{fig:video:rtt}
    \end{subfigure}%
    \caption{
        Multi-loop, resource-constrained setup results.
        Error bars indicate \SI{95}{\percent} \acp{CI} in all plots.
    }\label{fig:video:results}
\end{figure}

>>>>

Two main arguments drive this approach.
One, the difficulty of scaling the number of clients in these systems in a research context.
The systems we target exhibit a centralized nature in which a potentially large number of clients offload computation to a single central compute node.
The client-side component of the applications of interest for this research is often complex to scale, either due to the involvement of humans (as in the case of \gls{WCA}) or due its cyber-physical nature (in \glspl{NCS}).
Emulating this component reduces this complexity by moving it into the software domain, allowing for easier scaling through the use of cheap, general-purpose hardware such as single-board computers (e.g. Raspberry Pis, Beaglebones, etc.).

@@

In the following, we will discuss the initial implementation of this methodology to two different classes of applications.
We first discuss the methodology in the context of a human-in-the-loop application, \gls{WCA}, in section \cref{ssec:methodology:wca}.
Then, in \cref{ssec:methodology:ncs}, we present an exploration into it's utility for the study of \glspl{NCS}.

\subsection{Applying the methodology to \gls{WCA}}\label{ssec:methodology:wca}

@@

\subsection{Applying the methodology to \glspl{NCS}}\label{ssec:methodology:ncs}

@@

\begin{figure*}
    \centering
    \includegraphics[width=.8\textwidth]{publications/2022CLEAVE/images/CLEAVE_NCS_structure}
    \caption{
        Structure of an emulated \glspl{NCS} in \gls{CLEAVE}.
    }\label{fig:cleave:ncs:struct}
\end{figure*}

We overcome this issue by proposing a completely virtual plant allowing for unparalleled flexibility in changing the plant model and characteristic features of the experiments.
In this paper, we present the first fully-software-based framework for scalable and repeatable benchmarking of edge-native \gls{NCS}.
As edge computing begins being adopted by industry, more and more variations have begun to appear in literature.
``Near'', ``far'', ``core'', and ``telco'' edge describe variations of the original concept and are becoming ubiquitous in new research.
While the core idea of edge computing is widely accepted as fundamental for pervasive \glspl{NCS} in general, understanding the strengths and weaknesses of such different edge concepts is of paramount importance.

Our framework, \gls{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems.
It is fully virtualized, inspired by our previous work on benchmarking human-in-the-loop applications on the edge~\cite{Olguin2019EdgeDroid}.
The tool consists of a benchmarking framework and software development kit for the development of emulated physical systems and softwarized controllers.
These virtual \glspl{NCS} can then be deployed on real networks for reparametrizable, repeatable, and reproducible benchmarking of the complete system.

@@

We validate the utility of this tool through an example use case approximating a proposed edge deployment of inverted pendula control loops co-located with video analytics services.
We argue such a use case represents a realistic scenario and appropriate benchmark for the tool, since
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item the inverted pendulum plant is ubiquitous in \gls{NCS} research
    \item similar setups exist in real-world industrial use
    \item video analytics has long been proposed as a ``killer app'' for edge computing.
\end{enumerate*}
Our results showcase the ability of the framework to extract relevant metrics relating to the stability of the control system, as well as on the performance of the underlying network link.
We believe \gls{CLEAVE} represents an important step towards enabling inexpensive and low-complexity scalable research for real-world deployment of edge-bound \glspl{NCS}.

@@

\begin{figure}[t]
    \centering
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_topple_frac}
        \caption{Toppled plants.}\label{fig:video:toppled}
    \end{subfigure}%
    \hfill%
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_angle_rms}
        \caption{Angle \gls{RMS}.}\label{fig:video:rms}
    \end{subfigure}\\
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_drop_frac}
        \caption{Packet losses.}\label{fig:video:drop}
    \end{subfigure}%
    \hfill%
    \begin{subfigure}[h]{.22\textwidth}
        \centering
        \includegraphics[width=\textwidth]{publications/2022CLEAVE/plots/fixed_video_rtt}
        \caption{\glspl{RTT}.}\label{fig:video:rtt}
    \end{subfigure}%
    \caption{
        Multi-loop, resource-constrained setup results.
        Error bars indicate \SI{95}{\percent} \glspl{CI} in all plots.
    }\label{fig:video:results}
\end{figure}

### 2022-12-19 publications/2018DemoScalingOnTheEdge/paper_body.tex
glossaries
glossary fixes

    \item[Compute Intensive:] Cognitive Assistance applications aim to enhance the cognitive capabilities of users, and are thus compute intensive as well due to widespread use of the state-of-art computer vision and machine learning algorithms, particularly Deep Neural Networks (DNNs). 
    Although mobile devices are becoming increasingly powerful, the gap between mobile and static elements continues to exist~\cite{Flinn2012Cyber}. 
    While state-of-the-art DNN object detectors can run at more than 20 FPS on a server GPU, their performances are much worse on mobile GPUs - some models cannot even be loaded due to memory constraints. 
    Cloudlet-based applications overcome these challenges by offloading the computation. 
%    Offloading makes sense when the computation is intensive or special hardwares are needed.

@@

Some open questions and challenges remain to be tackled in the design and development of the presented tool.
To start with, the benchmarking suite is relatively narrow in the types of applications it can be applied to, currently only targeting event based cognitive assistance applications.
The tool needs to be extended to work on a much a broader spectrum of applications, in particular those that do not have a clear task model.
Furthermore, the current implementation does not support hardware accelerators (e.g. GPUs) that are commonly used for DNN inference.

>>>>

    \item[Compute Intensive:] Cognitive Assistance applications aim to enhance the cognitive capabilities of users, and are thus compute intensive as well due to widespread use of the state-of-art computer vision and machine learning algorithms, particularly \glspl{DNN}.
    Although mobile devices are becoming increasingly powerful, the gap between mobile and static elements continues to exist~\cite{Flinn2012Cyber}. 
    While state-of-the-art DNN object detectors can run at more than 20 \gls{FPS} on a server \gls{GPU}, their performances are much worse on mobile \glspl{GPU} - some models cannot even be loaded due to memory constraints.
    Cloudlet-based applications overcome these challenges by offloading the computation. 
%    Offloading makes sense when the computation is intensive or special hardwares are needed.

@@

Some open questions and challenges remain to be tackled in the design and development of the presented tool.
To start with, the benchmarking suite is relatively narrow in the types of applications it can be applied to, currently only targeting event based cognitive assistance applications.
The tool needs to be extended to work on a much a broader spectrum of applications, in particular those that do not have a clear task model.
Furthermore, the current implementation does not support hardware accelerators (e.g. \glspl{GPU}) that are commonly used for \gls{DNN} inference.

### 2022-12-19 publications/2022Ainur/abstract.tex
glossary fixes

\begin{abstract}
  Experimental research on wireless networking in combination with edge and cloud computing has been the subject of explosive interest in the last decade.
  This development has been driven by the increasing complexity of modern wireless technologies and the  extensive softwarization of these through projects such as a \acf*{O-RAN}.
  In this context, a number of small- to mid-scale testbeds have emerged, employing a variety of technologies to target a wide array of use-cases and scenarios in the context of novel mobile communication technologies such as 5G and beyond-5G.
  Little work, however, has yet been devoted to developing a standard framework for wireless testbed automation which is hardware-agnostic and compatible with edge- and cloud-native technologies.
  Such a solution would simplify the development of new testbeds by completely or partially removing the requirement for custom management and orchestration software.

  In this paper, we present the first such mostly hardware-agnostic wireless testbed automation framework, \emph{Ainur}.
  It is designed to configure, manage, orchestrate, and deploy workloads from an end-to-end perspective.
  Ainur is built on top of cloud-native technologies such as Docker, and is provided as \acs*{FOSS} to the community through the {KTH-EXPECA/Ainur} repository on GitHub.
  We demonstrate the utility of the platform with a series of scenarios, showcasing in particular its flexibility with respect to physical link definition, computation placement, and automation of arbitrarily complex experimental scenarios.
\end{abstract}

>>>>

\begin{abstract}
  Experimental research on wireless networking in combination with edge and cloud computing has been the subject of explosive interest in the last decade.
  This development has been driven by the increasing complexity of modern wireless technologies and the  extensive softwarization of these through projects such as a \gls{O-RAN}.
  In this context, a number of small- to mid-scale testbeds have emerged, employing a variety of technologies to target a wide array of use-cases and scenarios in the context of novel mobile communication technologies such as 5G and beyond-5G.
  Little work, however, has yet been devoted to developing a standard framework for wireless testbed automation which is hardware-agnostic and compatible with edge- and cloud-native technologies.
  Such a solution would simplify the development of new testbeds by completely or partially removing the requirement for custom management and orchestration software.

  In this paper, we present the first such mostly hardware-agnostic wireless testbed automation framework, \emph{Ainur}.
  It is designed to configure, manage, orchestrate, and deploy workloads from an end-to-end perspective.
  Ainur is built on top of cloud-native technologies such as Docker, and is provided as \gls{FOSS} to the community through the {KTH-EXPECA/Ainur} repository on GitHub.
  We demonstrate the utility of the platform with a series of scenarios, showcasing in particular its flexibility with respect to physical link definition, computation placement, and automation of arbitrarily complex experimental scenarios.
\end{abstract}

### 2022-12-19 publications/2022Ainur/ainur.tex
glossary fixes

    \item[\gls{OAI}:]
    \gls{OAI} is an open-source project that implements 3GPP technology on general purpose \texttt{x86} computing hardware and off-the-shelf \acp{SDR} like the \gls{USRP}~\cite{KALTENBERGER2020107284}.
    Ainur can deploy, configure, and manage \gls{OAI} software components that implement 4G \gls{LTE} and 5G \gls{NR}.

    \item[Mango Communications 802.11:]
    Finally, the Mango Communications project implements real-time 802.11 (WiFi) MAC and PHY in Xilinx \acp{FPGA}.
    It can be used on a variety of hardware platforms including \gls{USRP} \acp{SDR}, and is employed in Ainur to provision end-to-end WiFi links.
\end{description}

@@

Ainur follows a layered architecture which closely mimics the conceptual layers of the \acs{TCP}/\acs{IP} stack.
Components are deployed in bottom-up order, starting with the establishment of physical links, through the establishment of \gls{IP} connectivity and deployment of links to the cloud, and ending with the distribution and initialization of workloads on top of a container orchestration layer.
The architectural modules of the framework can broadly be classified according to the below categories:

@@

    \item[Physical Layer:] 
    Creates and deploys the underlying physical connections of the workload data network.
    This layer interacts with hardware such as managed switches and \acp{SDR} (and their associated computation hosts) to create links between the desired workload hosts.
    These links correspond to ethernet, WiFi, and even an entire software-defined 4G \gls{LTE} and 5G networks.

@@

As expressed above, Ainur also leverages containers for the communication stack.
With the advent of \gls{O-RAN} and \acp{SDR}, all components of the wireless network stack can run on general-purpose processors.
They can thus be containerized and distributed across multiple hosts.

@@

The framework is released as \acl{FOSS} under a permissive Apache license.
It can be downloaded from the Ainur repository~\cite{ainur:github} of the {KTH-EXPECA} organization on GitHub.
Given the software's complexity, the repository also includes detailed documentation on its requirements, configuration, deployment, and execution, as well as links to external resources containing information and guides about related concepts and services.

>>>>

    \item[\gls{OAI}:]
    \gls{OAI} is an open-source project that implements 3GPP technology on general purpose \texttt{x86} computing hardware and off-the-shelf \glspl{SDR} like the \gls{USRP}~\cite{KALTENBERGER2020107284}.
    Ainur can deploy, configure, and manage \gls{OAI} software components that implement 4G \gls{LTE} and 5G \gls{NR}.

    \item[Mango Communications 802.11:]
    Finally, the Mango Communications project implements real-time 802.11 (WiFi) MAC and PHY in Xilinx \glspl{FPGA}.
    It can be used on a variety of hardware platforms including \gls{USRP} \glspl{SDR}, and is employed in Ainur to provision end-to-end WiFi links.
\end{description}

@@

Ainur follows a layered architecture which closely mimics the conceptual layers of the \gls{TCP}/\gls{IP} stack.
Components are deployed in bottom-up order, starting with the establishment of physical links, through the establishment of \gls{IP} connectivity and deployment of links to the cloud, and ending with the distribution and initialization of workloads on top of a container orchestration layer.
The architectural modules of the framework can broadly be classified according to the below categories:

@@

    \item[Physical Layer:] 
    Creates and deploys the underlying physical connections of the workload data network.
    This layer interacts with hardware such as managed switches and \glspl{SDR} (and their associated computation hosts) to create links between the desired workload hosts.
    These links correspond to ethernet, WiFi, and even an entire software-defined 4G \gls{LTE} and 5G networks.

@@

As expressed above, Ainur also leverages containers for the communication stack.
With the advent of \gls{O-RAN} and \glspl{SDR}, all components of the wireless network stack can run on general-purpose processors.
They can thus be containerized and distributed across multiple hosts.

@@

The framework is released as \gls{FOSS} under a permissive Apache license.
It can be downloaded from the Ainur repository~\cite{ainur:github} of the {KTH-EXPECA} organization on GitHub.
Given the software's complexity, the repository also includes detailed documentation on its requirements, configuration, deployment, and execution, as well as links to external resources containing information and guides about related concepts and services.

### 2022-12-19 publications/2022Ainur/assumptions.tex
glossary fixes

In particular, the network follows a split architecture composed of two sub-networks: a control network for control plane data, and a workload data network for workload payload traffic.
It is designed like this to
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item be able to dynamically reconfigure, segment, and route the workload data network for different experiments
    \item ensure consistent and accurate network traffic measurements (\acp{RTT}, bandwidth, etc.) on the workload data network, uncontaminated by control traffic.
\end{enumerate*}

@@

    Devices on the workload data network are interconnected through a managed switch.
    This switch is used by Ainur to change the virtual topology of the workload data network at runtime using \acp{VLAN}, for instance separating the network into front-end and back-end networks.

@@

    \item[\acp{AMI}]

>>>>

In particular, the network follows a split architecture composed of two sub-networks: a control network for control plane data, and a workload data network for workload payload traffic.
It is designed like this to
\begin{enumerate*}[itemjoin={{, }}, itemjoin*={{, and }}]
    \item be able to dynamically reconfigure, segment, and route the workload data network for different experiments
    \item ensure consistent and accurate network traffic measurements (\glspl{RTT}, bandwidth, etc.) on the workload data network, uncontaminated by control traffic.
\end{enumerate*}

@@

    Devices on the workload data network are interconnected through a managed switch.
    This switch is used by Ainur to change the virtual topology of the workload data network at runtime using \glspl{VLAN}, for instance separating the network into front-end and back-end networks.

@@

    \item[\glspl{AMI}]

### 2022-12-19 publications/2022Ainur/demo.tex
glossary fixes

\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[height=8em]{publications/2022Ainur/figures/demo_workload_1.png}
        \caption{Workload 1: \acs*{MNIST} image classifier}\label{fig:wkld:mnist}
    \end{subfigure}%
    \vspace{1em}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[height=8em]{publications/2022Ainur/figures/demo_workload_2.png}
        \caption{Workload 2: inverted pendulum \acs*{NCS}}\label{fig:wkld:ncs}
    \end{subfigure}
    \caption{Demonstration workloads}\label{fig:wkld}
\end{figure}

@@

\Cref{fig:democonfigs} illustrates the possible configurations for the testbed.
Workloads are deployed on client hosts, and computation is offloaded either to an edge server (cloudlet), or to \gls{AWS} \gls{EC2} instances on the cloud.
Communication between the client- and server-sides of the workload will occur over one of three possible physical link layer setups:
\begin{description}
    \item[WiFi:] \iac{SDR} is configured as an IEEE 802.11n access point, and client hosts connect to it using on-board WiFi.
    \item[4G \gls{LTE}:] \iac{SDR} is configured as an 4G \gls{LTE} base station, and another is configured as an 4G \gls{LTE} \gls{UE}.
    Together these radios act as an \gls{LTE} bridge between the client-side and the server-side of the network, and all client-server traffic is routed through them.
    \item[Ethernet:] connects everything through plain ethernet.
\end{description}

@@

We will employ two different workloads for this demonstration, illustrated in \cref{fig:wkld}.
The first of these (\cref{fig:wkld:mnist}) consists of a proof-of-concept web application which implements a simple classifier to identify hand-drawn digits from the widely-used \gls{MNIST} dataset~\cite{mnist}.
The application consists of a client with a web interface, and \iac{HTTP} server which responds to requests from the client and performs the actual image recognition.
This workload is intended to showcase in an interactive manner the effects of placing computation at different points of the network, and how Ainur simplifies these deployments.

The second workload (\cref{fig:wkld:ncs}) corresponds to \iac{NCS} balancing an inverted pendulum, implemented on a software framework for the emulation of \acp{NCS} using cloud-native technologies~\cite{cleave}.
It consists of an emulation of the physical inverted pendulum system plant and a software-implemented proportional-differential controller.
These components communicate with each other over the \gls{UDP}.
This workload will be used to showcase the utility of Ainur for automating the execution of batches of experiments potentially including multiple different clients, servers, and physical layers.

@@

This demonstration will be performed on a testbed consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and finally }}]
    \item \num{10} Raspberry Pi 4 Model B boards, acting as client-side workload hosts
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the workload network
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the management network, as well as hosting the necessary \gls{NTP} and \gls{DNS} server software
    \item a \texttt{i386} workstation, acting as an edge-side workload host
    \item a configurable number of  \gls{AWS} \gls{EC2} cloud instances acting as cloud servers
    \item a separate \texttt{i386} workstation hosting the Fluent server and on which Ainur is deployed as well.
\end{enumerate*}
These nodes are interconnected using a combination of managed switches, \acp{SDR}, and \gls{VPN} gateways; please refer to \cref{fig:network} for an architectural overview of this setup.

@@

Participants will be asked to specify the full, arbitrarily complex, experimental scenario, including the number of clients, physical layers (single or multiple, and which clients are on each physical link), and offloading configuration (number of instances on the edge vs.\ on the cloud, which \gls{AWS} datacenters to deploy to).
Configuration will be specified through a \acs{YAML} file, which will then be parsed by Ainur for the automatic execution of the scenario.

>>>>

\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[height=8em]{publications/2022Ainur/figures/demo_workload_1.png}
        \caption{Workload 1: \gls{MNIST} image classifier}\label{fig:wkld:mnist}
    \end{subfigure}%
    \vspace{1em}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[height=8em]{publications/2022Ainur/figures/demo_workload_2.png}
        \caption{Workload 2: inverted pendulum \gls{NCS}}\label{fig:wkld:ncs}
    \end{subfigure}
    \caption{Demonstration workloads}\label{fig:wkld}
\end{figure}

@@

\Cref{fig:democonfigs} illustrates the possible configurations for the testbed.
Workloads are deployed on client hosts, and computation is offloaded either to an edge server (cloudlet), or to \gls{AWS} \gls{EC2} instances on the cloud.
Communication between the client- and server-sides of the workload will occur over one of three possible physical link layer setups:
\begin{description}
    \item[WiFi:] \pgls{SDR} is configured as an IEEE 802.11n access point, and client hosts connect to it using on-board WiFi.
    \item[4G \gls{LTE}:] \pgls{SDR} is configured as an 4G \gls{LTE} base station, and another is configured as an 4G \gls{LTE} \gls{UE}.
    Together these radios act as an \gls{LTE} bridge between the client-side and the server-side of the network, and all client-server traffic is routed through them.
    \item[Ethernet:] connects everything through plain ethernet.
\end{description}

@@

We will employ two different workloads for this demonstration, illustrated in \cref{fig:wkld}.
The first of these (\cref{fig:wkld:mnist}) consists of a proof-of-concept web application which implements a simple classifier to identify hand-drawn digits from the widely-used \gls{MNIST} dataset~\cite{mnist}.
The application consists of a client with a web interface, and \pgls{HTTP} server which responds to requests from the client and performs the actual image recognition.
This workload is intended to showcase in an interactive manner the effects of placing computation at different points of the network, and how Ainur simplifies these deployments.

The second workload (\cref{fig:wkld:ncs}) corresponds to \pgls{NCS} balancing an inverted pendulum, implemented on a software framework for the emulation of \glspl{NCS} using cloud-native technologies~\cite{cleave}.
It consists of an emulation of the physical inverted pendulum system plant and a software-implemented proportional-differential controller.
These components communicate with each other over the \gls{UDP}.
This workload will be used to showcase the utility of Ainur for automating the execution of batches of experiments potentially including multiple different clients, servers, and physical layers.

@@

This demonstration will be performed on a testbed consisting of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and finally }}]
    \item \num{10} Raspberry Pi 4 Model B boards, acting as client-side workload hosts
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the workload network
    \item a Raspberry Pi 4 Model B acting as \gls{VPN} gateway for the management network, as well as hosting the necessary \gls{NTP} and \gls{DNS} server software
    \item a \texttt{i386} workstation, acting as an edge-side workload host
    \item a configurable number of  \gls{AWS} \gls{EC2} cloud instances acting as cloud servers
    \item a separate \texttt{i386} workstation hosting the Fluent server and on which Ainur is deployed as well.
\end{enumerate*}
These nodes are interconnected using a combination of managed switches, \glspl{SDR}, and \gls{VPN} gateways; please refer to \cref{fig:network} for an architectural overview of this setup.

@@

Participants will be asked to specify the full, arbitrarily complex, experimental scenario, including the number of clients, physical layers (single or multiple, and which clients are on each physical link), and offloading configuration (number of instances on the edge vs.\ on the cloud, which \gls{AWS} datacenters to deploy to).
Configuration will be specified through a \gls{YAML} file, which will then be parsed by Ainur for the automatic execution of the scenario.

### 2022-12-19 publications/2022Ainur/futurework.tex
glossary fixes

    I (Manuel) suggest then that for ExPECA 2.0 we switch to a distributed client-server architecture, using a microservice for each core functionality.
    In other words, we would have several smaller server applications running on potentially different hosts on the network, each handling a specific part of the functionality we need.
    For instance, we could have a ``Layer 3'' service which handles setting up \gls{IP}-layer connectivity, and another ``SDR'' service which manages the \acp{SDR}.
    Services could communicate with each other using simple HTTP \acp{API} for easy debugging, and we would have an ``Ainur'' client which orchestrates the services into an experimental run.

>>>>

    I (Manuel) suggest then that for ExPECA 2.0 we switch to a distributed client-server architecture, using a microservice for each core functionality.
    In other words, we would have several smaller server applications running on potentially different hosts on the network, each handling a specific part of the functionality we need.
    For instance, we could have a ``Layer 3'' service which handles setting up \gls{IP}-layer connectivity, and another ``SDR'' service which manages the \glspl{SDR}.
    Services could communicate with each other using simple HTTP \glspl{API} for easy debugging, and we would have an ``Ainur'' client which orchestrates the services into an experimental run.
